Nitric oxide feedback to ciliary photoreceptor cells gates a UV avoidance circuit

### 2023-07-06 Manuscript/Manuscript.Rmd
MS: methods edit
Edit Figure 1 – figure supplement 1 legend
Ms: added iqtree command, updated bootstrap explanation

Through this optimisation procedure, we found parameters that gave a fit to our experimental data under all conditions.
We could fit to mean Ca^2+^ profiles but also individual Ca^2+^ transients (Figure 6B, Figure 6--figure supplement 1).
Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGWa in wild type and loss-of-function conditions (Figure 6A, B).
The model highlights that cPRC phototransduction employs distinct pathways that operate on different time scales and differentially influence cPRC Ca^2+^ levels.
The coupling between cPRCs and interneurons also requires different signalling mechanisms, either through distinct neurotransmitters or receptors expressed in the different cells.

@@

**Morpholino design**

@@

To identify NOS sequences, we obtained a seed database of oxygenase domains from the Pfam database (PF02898).
From these sequences, we generated a Hidden Markov Model (HMM) to mine the genomes and transcriptomes of 47 metazoan, 2 choanoflagellate and 2 filasterea species.
The HMM model was run in HMMR3 [@Eddy2011] with an e-value of 1e−15.
We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i [@Katoh2008].
The alignment was trimmed with TrimAl in gappy-out mode [@Capella-Gutierrez2009].
To calculate a maximum-likelihood tree, we used IQ-tree2 with the LG+G4 model.
To calculate branch support, we ran 1000 replicates with the ultrafast bootstrap approximation (UFBoot) [@Hoang2018] and aLRT-SH-like values [@Guindon2010] methods.

@@

To identify NIT-GCs, we obtained a seed database of adenylate and guanylate cyclase catalytic domains from the Pfam database (PF00211).
From these sequences, we generated a Hidden Markov Model (HMM) to mine the 45 metazoan, 2 choanoflagellate and 2 filasterea genomes and transcriptomes.
The HMM model was run in HMMR3 [@Eddy2011] with an e-value of 1e−15.
We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
To identify clusters, we used sequence-similarity-based clustering in Clans [@frickey2004] and the convex-clustering option with 100 jack-knife replicates.
The NIT-GCs are extremely well conserved within the membrane-bound guanylate cyclases and form an easily recognizable cluster.
To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building.
We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i [@Katoh2008].
The alignment was trimmed with TrimAl in gappy-out mode [@Capella-Gutierrez2009].
To calculate a maximum-likelihood tree, we used IQ-tree2 with the LG+G4 model.
To calculate branch support, we ran 1000 replicates with the ultrafast bootstrap approximation (UFBoot) [@Hoang2018] and aLRT-SH-like values [@Guindon2010] methods.

@@

```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 1. Maximum-likelihood phylogenetic tree of NOS protein sequences.** Individual branches are coloured by taxonomy. Support values are based on 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods. <br> Figure 1 -- figure supplement 1 -- source data 1. NOS sequences used for the phylogenetic reconstruction. <br> Figure 1 -- figure supplement 1 -- source data 2. Aligned and trimmed NOS sequences used for the phylogenetic reconstruction. <br> Figure 1 -- figure supplement 1 -- source data 3. Tree file of the reconstructed NOS phylogeny. "}
knitr::include_graphics(path = "figures/Fig1_sup1.png")

@@

```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 2. Maximum-likelihood phylogenetic tree of NIT-domain-containing guanylate cyclases** Membrane-bound and soluble guanylate cyclases (sGC) were included as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata. The expression of *Platynereis* NIT-GC genes in cPRC, INNOS and INRGW cells is indicated on the right side of the tree. The values represent expression in single-cell sequencing data. Dot size indicates specificity (percent of transcripts expressed in the indicated cell across all cells). Dot colour represents the logarithm of the number of transcripts in the expressing cells. <br> Figure 4 -- figure supplement 2 -- source data 1. GC sequences used for the phylogenetic reconstruction. <br> Figure 4 -- figure supplement 2 -- source data 2. Aligned and trimmed GC sequences used for the phylogenetic reconstruction. <br> Figure 4 -- figure supplement 3 -- source data 3. Tree file of the reconstructed GC phylogeny. "}
knitr::include_graphics(path = "figures/Fig4_sup2.png")

>>>>

Through this optimisation procedure, we found parameters that gave a fit to our experimental data under all conditions (Figure 6B; Figure 6--figure supplement 1-5).
We could fit to mean Ca^2+^ profiles but also individual Ca^2+^ transients.
Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGWa in wild type and loss-of-function conditions (Figure 6A, B).
The model highlights that cPRC phototransduction employs distinct pathways that operate on different time scales and differentially influence cPRC Ca^2+^ levels.
The coupling between cPRCs and interneurons also requires different signalling mechanisms, either through distinct neurotransmitters or receptors expressed in the different cells.

@@

**Morpholino-mediated knockdowns**

@@

To identify NOS sequences, we obtained a seed database of oxygenase domains from the Pfam database (PF02898).
From these sequences, we generated a Hidden Markov Model (HMM) to mine the genomes and transcriptomes of 47 metazoan, 2 choanoflagellate and 2 filasterea species.
The HMM model was run in HMMR3 [@Eddy2011] with an e-value of 1e−15.
We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i [@Katoh2008].
The alignment was trimmed with TrimAl in gappy-out mode [@Capella-Gutierrez2009].
To calculate a maximum-likelihood tree, we used IQ-tree2 with the LG+G4 model with the following options: iqtree2 -s "FILENAME" -mset WAG, LG, Blosum62, Dayhoff, JTT, Poisson -B 1000 -alrt 1000 -T AUTO.
Branch-support values are based on 1000 replicates with the ultrafast bootstrap approximation (UFBoot) [@Hoang2018] and aLRT-SH-like [@Guindon2010] methods.

@@

To identify NIT-GCs, we obtained a seed database of adenylate and guanylate cyclase catalytic domains from the Pfam database (PF00211).
From these sequences, we generated a Hidden Markov Model (HMM) to mine the 45 metazoan, 2 choanoflagellate and 2 filasterea genomes and transcriptomes.
The HMM model was run in HMMR3 [@Eddy2011] with an e-value of 1e−15.
We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
To identify clusters, we used sequence-similarity-based clustering in Clans [@frickey2004] and the convex-clustering option with 100 jack-knife replicates.
The NIT-GCs are extremely well conserved within the membrane-bound guanylate cyclases and form an easily recognizable cluster.
To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building.
We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i [@Katoh2008].
The alignment was trimmed with TrimAl in gappy-out mode [@Capella-Gutierrez2009].
To calculate a maximum-likelihood tree, we used IQ-tree2 with the LG+G4 model with the following options: iqtree2 -s "FILENAME" -mset WAG, LG, Blosum62, Dayhoff, JTT, Poisson -B 1000 -alrt 1000 -T AUTO.
Branch-support values are based on 1000 replicates with the ultrafast bootstrap approximation (UFBoot) [@Hoang2018] and aLRT-SH-like [@Guindon2010] methods.

@@

```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 1. Maximum-likelihood phylogenetic tree of NOS protein sequences.** Individual branches are coloured by taxonomy. Branch support values indicate UFBoot and aLRT-SH-like values. <br> Figure 1 -- figure supplement 1 -- source data 1. NOS sequences used for the phylogenetic reconstruction. <br> Figure 1 -- figure supplement 1 -- source data 2. Aligned and trimmed NOS sequences used for the phylogenetic reconstruction. <br> Figure 1 -- figure supplement 1 -- source data 3. Tree file of the reconstructed NOS phylogeny. "}
knitr::include_graphics(path = "figures/Fig1_sup1.png")

@@

```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 2. Maximum-likelihood phylogenetic tree of NIT-domain-containing guanylate cyclases** Membrane-bound and soluble guanylate cyclases (sGC) were included as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata. Branch support values indicate UFBoot and aLRT-SH-like values. The expression of *Platynereis* NIT-GC genes in cPRC, INNOS and INRGW cells is indicated on the right side of the tree. The values represent expression based on single-cell sequencing data. Dot size indicates specificity (percent of transcripts expressed in the indicated cell across all cells). Dot colour represents the logarithm of the number of transcripts in the expressing cells. <br> Figure 4 -- figure supplement 2 -- source data 1. GC sequences used for the phylogenetic reconstruction. <br> Figure 4 -- figure supplement 2 -- source data 2. Aligned and trimmed GC sequences used for the phylogenetic reconstruction. <br> Figure 4 -- figure supplement 3 -- source data 3. Tree file of the reconstructed GC phylogeny. "}
knitr::include_graphics(path = "figures/Fig4_sup2.png")
